---
title: "w41_tweets_facts_sentiment_FN"
author: "Arriaga, Hue, Khural, Moran"
date: "July 31, 2021"
output: html_document
---

```{r setup, include=FALSE}
library(data.table)
library(foreign)
library(tidyverse)
library(dplyr)
library(readr)


# Set working directory
setwd("C:/Users/Zaptetra/Google Drive/MIDS/W241/w241_final_project/notebooks")
knitr::opts_chunk$set(echo = TRUE)
```

## The raw data

First we will load the data.

```{r load data}
# d<- data.table::fread('../data/raw/response_tp_07_31.csv')
# d<- data.table::fread('../data/raw/response_tn_07_31.csv')
# d<- data.table::fread('../data/raw/response_fp_07_31.csv')
d<- data.table::fread('../data/raw/response_fn_07_31.csv')


```

## Data cleaning

We will first review:

- Eliminate empty columns
- Rename columns
- Update fields that didn't export correctly (missing data)
- Check for attrition or duplicates
- Check for fake answers: response time too quick plus attention check question wrong.
- Convert data to correct type (all are strings)

```{r pressure, echo=FALSE}

# delete unwanted rows
unwanted_rows <- c(1,2)

# # Drop blank columns
drop_columns <- c("RecipientLastName",
                   "RecipientFirstName",
                   "RecipientEmail",
                   "ExternalReference", "Progress", "Status",
                   "LocationLatitude",
                   "LocationLongitude","DistributionChannel",
                   "UserLanguage","IPAddress")

drop_col_index <- which(colnames(d) %in% drop_columns)

d<- d[-unwanted_rows, ]

d[, drop_columns]<-list(NULL)

col_new_names <- c("start_date", "end_date", "survey_duration",
                   "finished", "finished_date", "id", "email",
                   "gender", "age", "education",
                   "tweet_first_click_time", "tweet_last_click_time", "tweet_submit_time", "tweet_click_count",
                   "math_q1", "math_q2",
                   "test_first_click_time", "test_last_click_time", "test_submit_time", "test_click_count",
                   "stimulus","georgians","energy","soccer","fauci", "pollution","election")

# Rename columns
colnames(d) <- col_new_names

# Check for missing data
# Audit missing values
####################
# Survey id: R_3s64ZlPIAtROugC - missing soccer and fauci (response in qualtrics confirmed)
# d[id=="R_3s64ZlPIAtROugC"]$soccer <- "None of the above"
# d[id=="R_3s64ZlPIAtROugC"]$fauci <- "None of the above"

# survey id: R_33vRD5yKc2c3FIk - missing energy (response confirmed)
# d[id=="R_33vRD5yKc2c3FIk"]$energy <- "None of the above"

# # survey id: R_3j7jocPTJApWZXu - missing fauci
# d[id=="R_3j7jocPTJApWZXu"]$fauci <- "None of the above"

# survey id: R_2zoqcnPiyzpmJga - missing fauci
# d[id=="R_2zoqcnPiyzpmJga"]$fauci <- "None of the above"

# survey id: R_1I76YKkC3bSzfJv - missing fauci
# d[id=="R_1I76YKkC3bSzfJv"]$pollution <- "None of the above"
# d[id=="R_1I76YKkC3bSzfJv"]$georgians <- "None of the above"

# survey id: R_2asjDNe8PAuhxxh - missing fauci
# d[id=="R_2asjDNe8PAuhxxh"]$georgians <- "None of the above"

# survey id: R_d4o2OrwcVxrjEn7 - missing fauci
# d[id=="R_d4o2OrwcVxrjEn7"]$soccer <- "None of the above"

# survey id: R_8c7WtV8AzXTDACl - missing fauci
# d[id=="R_8c7WtV8AzXTDACl"]$soccer <- "None of the above"
####################
# After reviewing all the missing data, it seems that the option "None of the above" wasn't registered in any of the option

d[georgians=="" & finished=="True"]$georgians <- "None of the above"
d[energy=="" & finished=="True"]$energy <- "None of the above"
d[soccer=="" & finished=="True"]$soccer <- "None of the above"
d[fauci=="" & finished=="True"]$fauci <- "None of the above"
d[pollution=="" & finished=="True"]$pollution <- "None of the above"
d[election=="" & finished=="True"]$election <- "None of the above"

#duplicate rows
## ONLY FOR TN
## Respondent id-R_2dWlnlkYnG9epqt is a non-complier
# non_compliers_tn<-c("R_0IJ2yZ90ItKq2Fb")
# d<-d[!(id %in% non_compliers_tn),]

## ONLY FOR FP and FN
non_compliers_fp<-0
paste("We found",non_compliers_fp," non-compliers")



## Check for attrition
# People who received the treatment and then stopped
attrition_count <- length(d[finished=="False" & tweet_submit_time!=""]$id)

# People who received the treatment
total_responses <- length(d[tweet_submit_time!=""]$id)

attrition_proportion=attrition_count/total_responses
paste("We found a ", round(attrition_proportion*100, digits = 2),"% attrition.")

# Remove respondents that dropped (attrition + drop out)
respondent_drop <- which(d[,finished=="False"])
d <- d[-respondent_drop,]

# Save emails of raffle respondents
# write.csv(d$email,"../data/interim/tp_emails.csv", row.names = FALSE)
# write.csv(d$email,"../data/interim/tn_emails.csv", row.names = FALSE)
# write.csv(d$email,"../data/interim/fp_emails.csv", row.names = FALSE)
write.csv(d$email,"../data/interim/fn_emails.csv", row.names = FALSE)

# Data type conversion
d$start_date <- as.Date(d$start_date, format = "%Y-%m-%d %H:%M:%s")
d$end_date <- as.Date(d$end_date, format = "%Y-%m-%d %H:%M:%s")
# d$tweet_submit_time <- as.integer(d$tweet_submit_time)
# d$test_submit_time <- as.integer(d$test_submit_time)

```

We can also see a histogram of time looking at the tweets and time responding the test.

```{r}
par(mfrow=c(1,2))

h1 <- hist(d$tweet_submit_time,
     main = "Time reading Tweets",
     xlab = "seconds", 
     breaks = length(d$tweet_submit_time)/5)

h2<- hist(d$test_submit_time,
     main = "Time responding the test",
     xlab = "seconds", 
     breaks = length(d$test_submit_time)/5)


```

Data seems reasonably normal. Now let's remove variable that are not of interest.

```{r}
# Keep only columns of interest
colnames(d)
cols_to_remove <- c("start_date","end_date",
                    "finished","finished_date","id","email",
                    "tweet_first_click_time","tweet_last_click_time","tweet_click_count",
                    "test_first_click_time","test_last_click_time","test_click_count")

cols_to_drop_ix<-which(colnames(d) %in% cols_to_remove)

d[,cols_to_drop_ix]<- list(NULL)
```

Let's include the treatment columns to the data.

```{r}
## IMPORTANT Adding factors
# d$truth<- as.factor("fact")
d$truth<- as.factor("fake")
# d$sentiment<- as.factor("positive")
d$sentiment<- as.factor("negative")

# Correct answers for test: FACT
# d[ , bin_stimulus := ifelse(stimulus == "Green House Gases", yes = 1, no = 0)]
# d[ , bin_georgians := ifelse(georgians == "None of the above", yes = 1, no = 0)]
# d[ , bin_energy := ifelse(energy == "Natural Gas", yes = 1, no = 0)]
# d[ , bin_soccer := ifelse(soccer == "Unknown", yes = 1, no = 0)]
# d[ , bin_fauci := ifelse(fauci == "National Geographic Society", yes = 1, no = 0)]
# d[ , bin_pollution := ifelse(pollution == "6th largest", yes = 1, no = 0)]
# d[ , bin_election := ifelse(election == "2024", yes = 1, no = 0)]


# # Correct answers for test: FAKE
d[ , bin_stimulus := ifelse(stimulus == "Green House Gases", yes = 1, no = 0)]
d[ , bin_georgians := ifelse(georgians == "70", yes = 1, no = 0)]
d[ , bin_energy := ifelse(energy == "Wind", yes = 1, no = 0)]
d[ , bin_soccer := ifelse(soccer == "Pfizer vaccine", yes = 1, no = 0)]
d[ , bin_fauci := ifelse(fauci == "None of the above", yes = 1, no = 0)]
d[ , bin_pollution := ifelse(pollution == "1st largest", yes = 1, no = 0)]
d[ , bin_election := ifelse(election == "2024", yes = 1, no = 0)]

d$total_correct <- d%>% select(c("bin_stimulus", "bin_georgians","bin_energy","bin_soccer","bin_fauci","bin_pollution", "bin_election")) %>% rowSums()
```

We can see the distribution of correct answers.

```{r}
h3 <- hist(d$total_correct,
     main = "Correct answers distribution (max 7)",
     xlab = "correct answers", 
     breaks = length(d$total_correct)/4)
```

Now we can save the file for future reference and stacking.

```{r}
# Save to CSV
# write.csv(d,"../data/interim/tp_data.csv", row.names = FALSE)
# write.csv(d,"../data/interim/tn_data.csv", row.names = FALSE)
# write.csv(d,"../data/interim/fp_data.csv", row.names = FALSE)
write.csv(d,"../data/interim/fn_data.csv", row.names = FALSE)
```

Finally, we can stack all of the individual files.

```{r}
files <- c("tp_data.csv","tn_data.csv","fp_data.csv","fn_data.csv")

df_fn<-data.table::fread("../data/interim/fn_data.csv")
df_fp<-data.table::fread("../data/interim/fp_data.csv")
df_tn<-data.table::fread("../data/interim/tn_data.csv")
df_tp<-data.table::fread("../data/interim/tp_data.csv")

df<-rbind(df_tp,df_tn,df_fp,df_fn)

write.csv(df,"../data/processed/tweets_data.csv", row.names = FALSE)

```

Let's confirm data was saved properly

```{r}
df<-data.table::fread("../data/processed/tweets_data.csv")
df$truth_f <- as.factor(df$truth)
df$sentiment_f <- as.factor(df$sentiment)
```

